---
title: "German_rainfall"
author: "Emily Mendez"
date: "11/25/2025"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(zoo)
```


if date range is jan 02-31 chart doesnt include january to include it include 1 ; created new variable dates. 



```{r}
#Strings as Factors true
data23=read.csv("/Users/admin/Documents/Projects/Rainfall_Data.csv",stringsAsFactors = T)
data23


data23$dates=as.yearmon(paste(data23$Year, data23$Month), "%Y %m")
```

```{r}
sum(is.na(data23))

```

```{r}
head(data23)
```

```{r}
summary(data23)
```

```{r}
#Encode time variables
str(data23)
```

```{r}
data23$dates=as.yearmon(paste(data23$Year, data23$Month), "%Y %m")

```


```{r}

data23 %>% 
  group_by(City,Year) %>% 
  summarise(meanRainfall=mean(Rainfall..mm.))

```

```{r}
    #   msgs <- apply(messageData, 1, function(row) {
      #      messageItem(from = row[["from"]], message = row[["message"]])
       # })


input$SideBar_Names


   #box(width=12,
                         #   div(style="height:50px; font-size:20px;",
                          #      paste("Cities on the map:",paste(unique(data23$City),collapse = ", "))))

```


```{r}
pchMonth=substr(format(data23$dates, "%b"),1,2)
```

#222D32

```{r}

#by city and by year

berIndx=which(data23$City=="Berlin")

ggplot(data23[berIndx,])+
  aes(x=dates,Rainfall..mm.)+
  geom_line()+
  geom_text(aes(label =pchMonth[berIndx],color=(factor(pchMonth[berIndx]))),vjust=-1) 

```


```{r}
library(corrplot)
corData=cor(data23[,-c(1,8,11)])
corrplot(corData,method="circle",type="upper",
         title="Correlation Matrix",
         mar=c(0,0,2,0))
corData
```

```{r}
melty=melt(corData)
var1=melty[abs(melty$value)>0.6,]

indxw=attributes(var1)$row.names[which(var1$value<1)]

melty$Var2[indxw][1]

#var1

```

```{r}
knitr::kable(corData,format="markdown")
```


```{r}

library(leaflet)

leaflet(data23) %>% 
  addTiles() %>% 
  addMarkers(lng = data23$Longitude, lat=data23$Latitude, popup=data23$City)

```

`


```{r}
library(giscoR)
germany_counties <- gisco_get_nuts(
  year = "2021", #only 2016,2021, 2024 available, data from 2015-2023, most recent
  epsg = "4326", #gets coordinates
  resolution = "10",
  country = "DE", 
  nuts_level = "3" #regions
)
```



```{r}
Elevation_mm=data23$Elevation..m.
ggplot()+
  geom_sf(data=germany_counties,color = "black",fill="white")+
  geom_point(data=data23,size=5,shape = 16, fill = NA,
             aes(data23$Longitude,data23$Latitude,color=Elevation_mm))+
  scale_color_viridis_c()+
  coord_sf()+
  labs(title = "Elevation in 10 German Cities", x = "Longitude", y = "Latitude", fill = "Elevation (meters)")

```



```{r}
indicatorNames=colnames(data23[,-11])
indicatorInfo=c("Name of the city","City's latitude in degrees","City's longitude in degrees","The month number (1-12)",
                "The year of the data","Rainfall amount in millimeters","City’s elevation above sea level in meters",
                "The climate classification of the city", "Average temperature for the month in Celsius",
                "Average humidity level for the month in percentage")
indicator=cbind(indicatorNames,indicatorInfo)
colnames(indicator)=c("Variables","Description")

knitr::kable(indicator,format="markdown")


```





```{r}
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

point=st_as_sf(as.data.frame(data23[,c(2,3)]),coords = c("Longitude","Latitude"),crs=4326)
poinInter=st_transform(point,3035)
bounds=ne_countries(scale = "medium",returnclass = "sf")
germany=bounds[bounds$name=="Germany",]
```

```{r}
gridG=st_as_sf(grid, coords = c("Longitude", "Latitude"), crs = 4326)

inside <- lengths(st_within(gridG, germany)) > 0
gridG <- gridG[inside, ]

coords <- st_coordinates(gridG)
pred_df <- data.frame(Longitude = coords[,1], Latitude = coords[,2])



gridG$Elevation <- predict(gam_fit, newdata = pred_df)


```

```{r}
coordinates=as.data.frame(st_coordinates(gridG))

gridG$Longitude=coordinates$X
gridG$Latitude=coordinates$Y

ggplot() +
  geom_raster(data=gridG,aes(Longitude,Latitude,fill=Elevation)) +
  geom_sf(data=germany,color = "white",fill=NA)+
  geom_contour_filled()+
  scale_fill_viridis_c() +
  labs(title = "Elevation in Switzerland", x = "Longitude", y = "Latitude", fill = "Elevation (meters)")



```
```{r}


gam_fit <- gam(Elevation..m. ~ s(Longitude, Latitude, k = 10), data = data23)

lon.seq <- seq(st_bbox(bounds)['xmin'], st_bbox(bounds)['xmax'], length.out = 100)
lat.seq <- seq(st_bbox(bounds)['ymin'], st_bbox(bounds)['ymax'], length.out = 100)
grid <- expand.grid(Longitude = lon.seq, Latitude = lat.seq)

#grid$Elevation <- predict(gam_fit, newdata = grid)

```


```{r}


valueBox(paste(round(mean(data23$Temperature...C.),1),"\u00B0C"),
                                "Total Mean Temperature",icon=icon("temperature-low")),
                       valueBox(paste(round(mean(data23$Rainfall..mm.),1),"mm"),"Total Mean Rainfall",
                                icon=icon("cloud-showers-heavy")),
                       valueBox(round(mean(data23$Humidity....),1),
                                "Total Mean Humidity",icon=icon("cloud"))
                    ),


  fluidRow(
                         div(style = "text-align: center;",
                         actionButton("RenderRainPlot", "Go",icon=icon("play", style="color:black; opacity:0.8;"))
                     ))

```

```{r}
if ( ){
             output$RainfallCityPlot=renderPlot({
                 cityIndx=which(data23$City==input$CityChoose)
                 dataTemp=data23[cityIndx,] %>% 
                     filter(data23$dates>=as.Date(input$DateRange[1]),
                            data23$dates<=as.Date(input$DateRange[2]))
                 minyear=as.Date(input$DateRange[1])
                 minyr=format(minyear,"%Y")
                 maxyear=as.Date(input$DateRange[2])
                 maxyear=format(maxyear,"%Y")
                 
                 pchMonth=substr(format(dataTemp$dates, "%b"),1,2)
                 
                 ggplot(dataTemp)+
                     aes(x=dates,Rainfall..mm.,color=City)+
                     geom_line()
                     geom_text(aes(label = pchMonth[cityIndx],color=(factor(pchMonth[cityIndx]))),vjust=-1) +
                     ggtitle(paste0(City,"Rainfall over ",minyr,"-",maxyear))
             })}
```




```{r}
  output$RainfallCityPlot=renderPlot({
        pchMonth=substr(format(data23$dates, "%b"),1,2)
        data23$PchVal=pchMonth[1:length(data23$City)]
        
        Datasub=data23 %>% 
            filter(City=="Berlin")
        
        ggplot(Datasub)+
            aes(x=dates,Rainfall..mm.)+
            geom_line()+
            geom_text(aes(label = PchVal,color=(factor(PchVal))),vjust=-1)+
            ggtitle(paste0(Datasub$City[1]," Rainfall over ",min(Datasub$Year),"-",max(Datasub$Year)))
    })
    




  output$RainfallCityPlot=renderPlot({
        minyear=input$DateRange[1]
        minyr=format(minyear,"%Y")
        maxyear=input$DateRange[2]
        maxyear=format(maxyear,"%Y")
        pchMonth=substr(format(dataTemp$dates, "%b"),1,2)
        dataTemp$pchM=pchMonth[1:length(dataTemp$dates)]
            
        ggplot(dataTemp)+
            aes(x=dates,y=Rainfall..mm.)+
            geom_line()+
            geom_text(aes(label =pchM ,color=(factor(pchM))),vjust=-1) +
            ggtitle(paste0(City,"Rainfall over ",minyr,"-",maxyear)
                    
                    
                     dataTemp=data23[which(data23$City==input$CityChoose),] %>% 
            filter(data23$dates>=as.Date(input$DateRange[1]),
                   data23$dates<=as.Date(input$DateRange[2]))
        print(dataTemp)
```




  output$RainfallCityPlot=renderPlot({
            minyear=as.Date(input$DateRange[1])
            minyr=format(minyear,"%Y")
            maxyear=as.Date(input$DateRange[2])
            maxyear=format(maxyear,"%Y")
            pchMonth=substr(format(dataTemp$dates, "%b"),1,2)
            dataTemp$pchM=pchMonth[1:length(dataTemp$dates)]
            monthLetter=factor(pchM)
            
            ggplot(dataTemp)+
                aes(x=dates,y=Rainfall..mm.)+
                geom_line()+
                geom_text(aes(label =pchM ,color=(monthLetter),vjust=-1)) +
                ggtitle(paste0(dataTemp$City[1],"Rainfall over ",minyr,"-",maxyear))
            })

```{r}
  ggplot(data23[])+
        aes(x=dates,y=Rainfall..mm.)+           
        geom_line()+
        geom_text(aes(label =pchM ,color=(monthLetter),vjust=-1)) +
        ggtitle(paste0(dataTemp$City[1],"Rainfall over ",minyr,"-",maxyear))
```


```{r}
ggplot(data23[which(data23$City=="Berlin"),])+
  aes(Year,Rainfall..mm.)+
  geom_point()
```


```{r}
ggplot(data23)+
  aes(data23$Temperature...C.,data23$Humidity....)+
 # geom_errorbar(aes(ymin=mean(Temperature...C.)-sd(Temperature...C.), 
  #                  ymax=mean(Temperature...C.)+sd(Temperature...C.)))+
  geom_boxplot()   
  
```

temp: city, climate type
rain: city, climate 

```{r}
  outputcorrBox=renderInfoBox(
        melty=melt(corData),
        melty2=melty[1>abs(melty$value)>0.6,],
        valueBox(paste("The Variables with moderate to high correlation are: ",
                       melty2$Var1," and ", melty2$var2),
                           icon=icon("arrow-up-right-dots",style = "color:white; opacity:0.5;"),color="navy")
    )
    
    valueBox(round(mean(clicker$Humidity),1), paste("Total Avg. Humidity of",clicker$City[1]), 
             icon=icon("cloud",style = "color:white; opacity:0.5;"),color="navy")
```


```{r}

x=c()

for(i in colnames(data23)){
  if(!is.null(levels(data23[,i]))){
    x=c(x,i)
  }}

x[2]
```


```{r}
dataTempMat=data23 %>% 
  group_by(City) %>% 
  summarise(AvgTemp=mean(Rainfall..mm.))

dataTempMat
```


```{r}
ggplot(dataTempMat)+
          aes(City,AvgTemp,fill=City)+
          geom_bar(stat = "identity")


```




```{r}

 box(status = "primary",width = NULL,height=200,
                                 selectInput("",label = "Choose a Variable to model as the response:",
                                             choices = colnames(data23[,c(6,9,10)]))),
                             
                             box(status = "primary",height=200,
                                 width = NULL, selectInput("", label = "Choose Variables to use as predictors ",
                                                           choices = colnames(data23[,c(1,8)]))), 

```


```{r}

  
    output$DataChart=renderDT({
      if (is.null(input$predictor)){
        datatable(data23[,-11],options = list(lengthMenu = c(15, 20, 50, 70),scrollX=TRUE),
                  rownames = FALSE)
      }
      else{
        datatemp=data23[,-11]
        datatable(datatemp[,c(input$response,input$predictor)],options = list(lengthMenu = c(15, 20, 50, 70),scrollX=TRUE),
                  rownames = FALSE)
      }
    })

```


Model
Diagnostics
Prediction

```{r}
set.seed(123)
indx=createDataPartition(data23$Rainfall..mm.,times=1,p=.7,list=FALSE)
  train=data23[indx,-11]
  test=data23[-indx,-11]

```


```{r}
##Perfect mulicollinarity perfectly reconstruct the city dummies from the coordinates, which is why VIFs explode.
#then City is essentially a categorical encoding of location, while longitude/latitude (and often elevation) are continuous encodings of the same thing
lmoddr=lm(Rainfall..mm. ~.,data23[,-c(11,4,5)])
summary(lmoddr)
ols_vif_tol(lmoddr)
ali=alias(lmos)$Complete


lmos=lm(Rainfall..mm. ~ Humidity....,data23[,-c(11,4,5)])
rownames(ali);colnames(ali)



  rownames(ali)[,which(ali>1)]
  indx=)
        ,arr.ind = TRUE)


  indx
  rowq=rownames(ali)[indx[,1]]
  colq=colnames(ali)[indx[,2]]
  
  cbind(rowq,colq)
  table(cbind(rowq,colq))

  importance=as.data.frame(varImp(lmos))
  importance
  
 
  
  as.formula(paste("City",'~1'))
```


```{r}
ols_vif_tol(lmoddr)

data34$rrr=data34$Rainfall..mm.
data56=data34[,-c(7,3)]
colnames(data34)
lmod32=lm(Rainfall..mm.~Temperature...C.+Humidity....+Climate_Type,data=data34)

ols_step_forward_p(lmod32)
ols_step_both_p(lmoddr)


alias()



```

```{r}
data34=data23[,-c(11,1,4,5,8)] %>% 
  mutate_all(~(scale(.) %>% as.vector))

```

```{r}
  newPredictors = setdiff(input$predictor, input$response)
         preds=paste(newPredictors, collapse = " + ")
         lmodString=as.formula(paste(input$response,'~',preds))
         resps=as.formula(paste(input$response,'~1'))
         lmod1=lm(lmodString,data=train)
         summary(lmod1)
         
         
         
         lmod2=lm(as.formula(paste(input$response, "~.")),data = train)
         summary(lmod2)
         
         
         
          
        else{

          indx=which(abs(ali)>1,arr.ind = TRUE)
          rowq=rownames(ali)[indx[,1]]
          colq=colnames(ali)[indx[,2]]
          cat("Frequency of Redundant Variables")
          table(cbind(rowq,colq))
        }
```

```{r}
residsPlot
```

